Syllabus

### 2019-08-22 content/syllabus.Rmd
Cleared out old content, updated syllabus, reference (tips for success), and schedule based on devF19

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/files/syllabus.pdf)

@@

<span class="newthought">\textbf{Fair warning:}</span> **Econometrics is hard.** *It will be one of the hardest economics courses that you will take, primarily due to the mathematical content.* I will do my best to make this class intuitive and helpful, if not interesting. If at any point you find yourself struggling in this course for any reason, please come see me. Do not suffer in silence! Coming to see me for help does not diminish my view of you, in fact I will hold you in *higher* regard for understanding your own needs and taking charge of your own learning. There are also a some fantastic resources on campus, such as the [Center for Academic Achievement and Retention (CAAR)](http://www.hood.edu/campus-services/academic-services/index.html) and the [Beneficial-Hodson Library](http://www.hood.edu/library/}{Beneficial-Hodson Library).

@@

<span class="newthought">Most of the readings in this class are free.</span>

@@

- [@Bailey], 2^nd^ ed. ISBN: 978-0190857462

You are welcome to purchase the book by other means (e.g. [Amazon](https://www.amazon.com/Real-Econometrics-Answer-Important-Questions/dp/0190857463/ref=dp_ob_title_bk), half.com, etc). I have no financial stake in requiring you to purchase this book. The (cheaper) 1st edition is sufficient, but makes significantly less use of `R` (in favor of `STATA`).

@@

Throughout the course, I will post both required and supplemental (non-required) readings that enrich your understanding for each topic on Blackboard. Check Blackboard *frequently* for announcements and updates to assignments, readings, and grades.

## Online help and Slack

Economics can be difficult. Computer programming can be difficult. Computers are stupid and little errors in your code can cause hours of headache (even if you've been doing this stuff for years!).

Fortunately there are tons of online resources to help you with this. Two of the most important are [StackOverflow](https://stackoverflow.com/) (a Q&A site with hundreds of thousands of answers to all sorts of programming questions) and [RStudio Community](https://community.rstudio.com/) (a forum specifically designed for people using RStudio and the tidyverse (i.e. you)).

Searching for help with R on Google can be tricky because the program is, um, a single letter. Try searching for "rstats" instead. If you use Twitter, post R-related questions and content with [#rstats](https://twitter.com/search?q=%23rstats).

Additionally, we have a class chatroom at [Slack](https://hoodeconometrics.slack.com/) where anyone in the class can ask questions and anyone can answer. Ask questions about the readings or problem sets [in the class Slack workspace](https://hoodeconometrics.slack.com/). I will monitor Slack regularly, and you should also all do so as well. You'll likely have similar questions as your peers, and you'll likely be able to answer other peoples' questions too.

# Course policies

<span class="newthought">Be nice. Be honest. Don't cheat.</span>

We will also follow the [full list of Marriott School and BYU classroom policies](https://marriottschool.byu.edu/policies/classroom/).

## Counseling and Psychological Services (CAPS)

Life at BYU can be complicated and challenging. You might feel overwhelmed, experience anxiety or depression, or struggle with relationships or family responsibilities. [Counseling and Psychological Services (CAPS)](https://caps.byu.edu/) provides free, *confidential* support for students who are struggling with mental health and emotional challenges. The CAPS office is staffed by professional psychologists who are attuned to the needs of all types of college and professional students. Please do not hesitate to contact CAPS for assistance—getting help is a smart and courageous thing to do.

@@

```{r}
knitr::kable(assignments)
```

@@

\

@@

# Red pandas

[^panda] <span class="newthought">Once you have read</span> this entire syllabus and [the assignments page](/assignments/), please [click here](mailto:andrew_heiss@byu.edu) and e-mail me a picture of a red panda. For real. Brownie points if it's animated.

[^panda]: {-}
  ![Hungry red panda](https://media.giphy.com/media/r7Nbr97FsXxXG/giphy.gif)

>>>>

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/files/syllabus.pdf)
- **Last Updated**: 2019-08-22

@@

<span class="newthought">Fair warning:</span> **Econometrics is hard.** *It will be one of the hardest economics courses that you will take, primarily due to the mathematical content.* I will do my best to make this class intuitive and helpful, if not interesting. If at any point you find yourself struggling in this course for any reason, please come see me. Do not suffer in silence! Coming to see me for help does not diminish my view of you, in fact I will hold you in *higher* regard for understanding your own needs and taking charge of your own learning. There are also a some fantastic resources on campus, such as the [Center for Academic Achievement and Retention (CAAR)](http://www.hood.edu/campus-services/academic-services/index.html) and the [Beneficial-Hodson Library](http://www.hood.edu/library/}{Beneficial-Hodson Library).

See my [tips for success in this course](/reference#tips).

@@

@@

- Bailey, Michael A, 2019, *Real Econometrics*, New York: Oxford University Press, 2^nd^ ed.

You are welcome to purchase the book by other means (e.g. Amazon, half.com, etc). I have no financial stake in requiring you to purchase this book. The (cheaper) 1st edition is sufficient, but makes significantly less use of `R` (in favor of `STATA`).

The following two books are **recommended**, and are free online^[You can purchase a hard copy of the first one if you really want.]:

- Grolemund, Garrett and Hadley Wickham, [*R For Data Science*](https://r4ds.had.co.nz/)
- Ismay, Chester and Albert Y Kim, [*Modern Dive: Statistical Inference Via Data Science*](https://moderndive.com/)

The first book is the number one resource for using `R` and `tidyverse`, and is written for beginners. I still look at it *frequently.* The second is another great reference for using `tidyverse` in the context of basic statistics.

@@

Throughout the course, I will post both required and supplemental (non-required) readings that enrich your understanding for each topic. Check Blackboard *frequently* for announcements and updates to assignments, readings, and grades.

@@

```{r, echo=F}
knitr::kable(assignments)
```

@@

All grades are based on the following traditional scale:

@@

These grades are firm cutoffs, but I do of course round upwards ($\geq 0.5$) for final grades. A necessary reminder, as an academic, I am not in the business of *giving* out grades, I merely report the grade that you *earn*. I will not alter your grade unless you provide a reasonable argument that I am in error (which does happen from time to time).

# Policies and Expectations

This syllabus is a contract between you, the student, and me, your instructor. It has been carefully and deliberately thought out^[A syllabus can and will be used as a legal document for disputes tried at a court of law. Ask me how I know.], and I will uphold my end of the agreement and expect you to uphold yours.

In the language of game theory, this syllabus is my commitment device. I am a very understanding person, and I know that exceptions to rules often need to be made for students. However, to be *fair* to *all* students the syllabus artificially constrains my ability to make exceptions at a whim for anyone. This prevents clever students from exploiting my congenial personality at everyone else's expense. Please read and familiarize yourself with the course policies and expectations of you. Chances are, if you have a question, it is answered herein.

### Attendance and Participation

I expect you to attend class and to come having already done the reading assigned for that day. I will remind you in class and possibly through Blackboard or email which readings I want you to read for the next class. You are all adults and I will treat you as such. I do not take attendance, nor do I grade formally for participation but I strongly recommended you attend class and participate for your sake and the sake of your classmates. If you are too distracted or are not prepared to learn, I suggest you stay home, where you can check Facebook more efficiently. I reserve the right to boost the final grades of students that I believe have made consistent, quality contributions above and beyond their peers in class conversations by up to 2.5 points.

### Devices

You are allowed to have and use laptops and tablets in the classroom. I will not stop you, but I strongly discourage you from using these to take notes (see Tips for Success). As a courtesy to myself and to others, do not use your phone in class. I reserve the right to embarrass you in front of everyone if you do so.

### Absences and Make-Ups

You generally do *not* need to let me know if you are unable to make class, *unless* it is on the day of an exam. It will however, be your responsibility to acquire the notes from a classmate for any missed classes. If you are unable to attend an exam for a legitimate reason (e.g. sports/club events, traveling, illness, family issues), please notify me at least *one week* in advance, and we will schedule a make-up exam date. If you are ill or otherwise unable to attend on the day of the exam, contact me ASAP to make arrangements. Failure to do so, including desperate attempts to make arrangements only *after* the absence will result in a grade of 0 and little sympathy. I reserve the right to re-weight other assignments for students who I believe are legitimately unable to complete a particular assignment.

### Late Assignments

I will accept late assignments, but will subtract a specified amount of points as a penalty. See individual assignment descriptions for the amount of points taken off (as it varies by assignment). If an answer key is posted before you turn in your assignment, the maximum grade you can earn is an 80. Even if it is the last week of the semester, I encourage you to turn in late work: some points are better than no points!

### Grading

I will try my best to post grades on Blackboard's Grading Center and return graded assignments to you within about one week of you turning them in. There will be exceptions. Where applicable, I will post answer keys once I know most homeworks are turned in (see Late Assignments above for penalties). Blackboard's Grading Center is the place to look for your most up-to-date grades. You will also be given an Excel spreadsheet template where you can calculate your overall grade and forecast "what if" scenarios.

### Email Accounts

Students must regularly monitor their Hood email accounts to receive important college information, including messages related to this class. Email through the Blackboard system is my main method of communicating announcements and deadlines regarding your assignments. **Please do not reply to any automated Blackboard emails - I may not recieve it!**. My Hood email (safner@hood.edu) is the best means of contacting me. I will do my best to respond within 24 hours. If I do not reply within 48 hours, do not take it personally, and *feel free to send a follow up email* in the very likely event that I genuinely did not see your original message.

### Office Hours

I am generally in my office Monday-Thursday during "normal business hours." You are always welcome to walk-in and chat about class, college, careers, or anything at all. Please do try to use the official office hours stated at the head of the syllabus if possible. If you need to meet at a different time, I request that you send me an email or let me know after class so I know when to expect you. If you want to go over material from class, please have *specific* questions you want help with. I am not in the business of giving private lectures (particularly if you missed class without a valid excuse).

Watch this excellent and accurate video explaining office hours:

<iframe src="https://player.vimeo.com/video/270014784" width="640" height="360" frameborder="0" allow="autoplay; fullscreen" allowfullscreen></iframe>

### Van Halen and M&Ms

When you have completed reading the syllabus, email me a picture of the band Van Halen and a picture of a bowl of M&Ms. If you do this *before* the date of the first exam, you will get 3 bonus points on the exam. Yes, this is real.

### Enrollment

Students are responsible for verifying their enrollment in this class. The last day to add or drop this class with no penalty is **Tuesday, September 4**. Be aware of [important dates](https://www.hood.edu/offices-services/registrars-office/academic-calendar).

### Honor Code

Hood College has an Academic Honor Code which requires all members of this community to maintain the highest standards of academic honesty and integrity. Cheating, plagiarism, lying, and stealing are all prohibited.  All violations of the Honor Code are taken seriously, will be reported to appropriate authority, and may result in severe penalties, including expulsion from the college. See [here](http://hood.smartcatalogiq.com/en/2016-2017/Catalog/The-Spirit-of-Hood/The-Academic-Honor-Code-and-Code-of-Conduct) for more detailed information.

### Accessibility, Equity, and Accommodations

College courses can, and should, be challenging and bring you out of your comfort zone in a safe and equitable environment. If, however, you feel at any point in the semester that certain assignments or aspects of the course will be disproportionately uncomfortable or burdensome for you due to any factor beyond your control, please come see me or email me. I am a very understanding person and am happy to work out a solution together. I reserve the right to modify and reweight assignments at my sole discretion for students that I belive would legitimately be at a disadvantage, through no fault of their own, to complete them as described.

If you are unable to afford required textbooks or other resources for any reason, come see me and we can find a solution that works for you.

This course is intended to be accessible for all students, including those with mental, physical, or cognitive disabilities, illness, injuries, impairments, or any other condition that tends to negatively affect one's equal access to education. If at any point in the term, you find yourself not able to fully access the space, content, and experience of this course, you are welcome to contact me to discuss your specific needs. I also encourage you to contact the [Office of Accessibility Services](https://www.hood.edu/academics/josephine-steiner-center-academic-achievement-retention/accessibility-services) (301-696-3421). If you have a diagnosis or history of accommodations in high school or previous postsecondary institutions, Accessibility Services can help you document your needs and create an accommodation plan. By making a plan through Accessibility Services, you can ensure appropriate accommodations without disclosing your condition or diagnosis to course instructors.

# Schedule

<span class="newthought">You can find a full schedule</span> with resources for each class meeting on the [schedule page](/schedule/).
